\section{Conclusion}

We were successfully able to show the microscopic force-based pedestrian dynamics model through port-Hamiltonian formulation. We witnessed a shift in perspective from microscopic models, where we inspect agent interactivity on an individual level; to a holistic description of the system-wide interactivity through measuring the system energy. This energy-based perspective, enabled by port-Hamiltonian systems, allowed us to not only observe collective behavior, but also to identify it through quantitative means using the Hamiltonian and $H^*$. With the port-Hamiltonian formulation, we were able to replicate the findings from \cite{tordeux2022multi} of well-known collective phenomenon such as lane and stripe formation. 

We can also observe that the interaction between the pedestrians is isotropic, this is due to the skew-symmetric nature of Hamiltonian systems. A physical interpretation of this, in the context of our model, assumes that the pedestrians interact equally with all of its surroundings, hence there are no vision cones effects that may produce bias in interaction in the direction of motion \cite{helbing1995social}. An attempt to incorporate such anisotropic effects within the port-Hamiltonian approach includes the use of state-dependent input terms \cite{tordeux2022multi} though it remains an active part of research.

Furthermore, we were also able to show the impact of induced noise on these collective phenomenon, in particular, "noise-induced ordering" for stripe formation \cite{khelfa2021initiating,d2021canard}. In the case of lane formation, the collective phenomenon was disrupted with increase in volatility, which is in accordance with the "freezing by heating" phenomenon \cite{helbing2000freezing}, however formation of the clusters of agents could not be reproduced.

With this thesis, we've been able to show that port-Hamiltonian systems, although a relatively young area of research, promises a modeling approach to quantify collective dynamics. Because it models systems from the perspective of energy-based interactions, it is more abstract, and thus useful to understand the underlying dynamics of systems from to various disciplines.

Working on this project has been a great learning experience for me. The study of complex systems, collective behavior, emergent phenomenon, and nonlinear dynamics have always been alluring to me, yet also enigmatic. I am glad to have been introduced to a topic such as stochastic port-Hamiltonian systems as they provide a good tool for studying these topics. With this I've been introduced to a plethora of different paths to take, each a world of its own, such as stochastic differential equations, port-Hamiltonian systems on infinite-dimensional spaces, geometric integration, and differential forms, to name a few. These are a few topics that I would like to explore more as I dive deeper into the study of collective phenomenon through the lens of port-Hamiltonian systems.

Thank you.